Zohmg---a large scale data store for aggregated time-series-based data. [DRAFT]

### 2009-05-06 doc/report/msc-report.tex
doc/report/msc-report.tex: + changed "data warehouse" to "data store". + added citations to tools.
doc/report/references.bib:
+ added references to tools.
+ cosmetic changes.
doc/report/msc-report.tex: * changed "dataset" to "data set". * added text in sections about data server (implementation and usage). * added brief outline to web log analysis section. * added one sentence to conclusion.

\author{Per Andersson, Fredrik M{\"o}llerstrand}
\title{Zohmg---a large scale data warehouse for aggregated time-series-based
data. [DRAFT]}

@@

Scalability is also very important. Using single machines for computing is
problematic when the dataset grows, more powerful machines has to be
obtained. More power contained in a single machine results in higher price,
eventually it is too expensive to perform computations on very large
datasets. Both MapReduce and BigTable scale close to linearly, adding one
machine to the cluster adds equal amount of processing power. Since
MapReduce and BigTable can use commodity hardware it is reasonable to buy
extra machines to gain extra computing power or storage space.

@@

A typical projection is that from an n-dimensional cube down to two dimensions, which is the case when plotting an image on screen or paper.

@@

In actual practice, the main difference between the two models is that
querying is highly restricted in a BigTable-like data store---there are no
joins, for example---whereas the relational database was born to be queried.
On the other hand, the restrictions placed upon the BigTable models are
what allows it to scale to such a large dataset. Indeed, BigTable was born
to scale.

@@

Hadoop is a free software implementation of Google's MapReduce-framework
described in \cite{mapreduce}. Hadoop is written in Java and is part of
the Apache Software Foundation.  It is a fairly mature piece of software
and is used in production at numerous companies.

@@

HBase is a free software implementation of Google's BigTable data store
described in \cite{bigtable}. It is a sub-project of Hadoop, written in
Java and a part of the Apache Software Foundation. HBase is still in its
infancy and few organizations use it for mission-critical applications.

@@

Dumbo is a framework for writing Hadoop programs in languages other than Java, in a so-called Streaming Mode. Dumbo is written in Python. It is used extensively at Last.fm for writing short prototypes.

@@

Most Hadoop users write their MapReduce programs in Java, since this is the
native language of Hadoop.

@@

Python is a popular general-purpose high-level multi-paradigm---most
notably: object-oriented, imperative and functional---scripting language.

@@

Python Paste is a framework for building WSGI\footnote{Web Server Gateway
Interface, a HTTP extension where a CGI-like environment is passed around.}
applications and middleware.

@@

Middleware is software that acts as an intermediary. WSGI middleware
receives a WSGI request and then performs logic upon this request, before
passing the request to a WSGI application.

@@

PyYAML is a complete YAML 1.1 compliant parser and emitter for Python.

@@

Python Setuptools is a framework---built on top of the Python module
distutils---for basic package management of Python software. Amongst other
things it creates and installs Python Eggs\footnote{A Zip file which
bundles Python software files.}.

@@

A simple, fast, complete and extensible JSON encoder and decoder for
Python. The Python module simplejson is implemented entirely in Python and
has no external dependencies. However a C extension which provides a
serious speed boost is included.

@@

The user's datasets are defined using YAML. The dataset files reside in
the config directory.

@@

@@

\subsubsection{JSON}

@@

To extract data, visit the server URL (e.g. http://localhost:8086/) and
give a query request.

\subsection{Data server}

Built with Python Paste---a middleware framework---the data server
dispatches to the requested application depending on the URI it was fed.

Middleware which dispatches to applications based on the URI.

Dispatches request to serve static files from the project's client directory.
Custom clients can be built with HTML and JavaScript, see the examples.

Serves data from HBase as JSONP\footnote{JSON with padding: JSON extension which adds name of callback function as an argument to the function itself.}.

@@

@@

\subsection{Multiple datasets}

@@

\subsection{Updating dataset files}

Update tables on changes in configuration---dataset files.

@@

>>>>

\author{Per Andersson, Fredrik M{\"o}llerstrand}
\title{Zohmg---a large scale data store for aggregated time-series-based
data. [DRAFT]}

@@

Scalability is also very important. Using single machines for computing is
problematic when the data set grows, more powerful machines has to be
obtained. More power contained in a single machine results in higher price,
eventually it is too expensive to perform computations on very large
data sets. Both MapReduce and BigTable scale close to linearly, adding one
machine to the cluster adds equal amount of processing power. Since
MapReduce and BigTable can use commodity hardware it is reasonable to buy
extra machines to gain extra computing power or storage space.

@@

A typical projection is that from an n-dimensional cube down to two
dimensions, which is the case when plotting an object---with three or more
dimensions---on screen or paper.

@@

In actual practice, the main difference between the two models is that
querying is highly restricted in a BigTable-like data store---there are no
joins, for example---whereas the relational database was born to be queried.
On the other hand, the restrictions placed upon the BigTable models are
what allows it to scale to such a large data set. Indeed, BigTable was born
to scale.

@@

Hadoop is a free software implementation of Google's MapReduce-framework
described in \cite{mapreduce}. Hadoop is written in Java and is part of
the Apache Software Foundation.  It is a fairly mature piece of software
and is used in production at numerous companies. \cite{hadoop}

@@

HBase is a free software implementation of Google's BigTable data store
described in \cite{bigtable}. It is a sub-project of Hadoop, written in
Java and a part of the Apache Software Foundation. HBase is still in its
infancy and few organizations use it for mission-critical applications.
\cite{hbase}

@@

Dumbo is a framework for writing Hadoop programs in languages other than
Java, in a so-called Streaming Mode. Dumbo is written in Python. It is used
extensively at Last.fm for writing short prototypes. \cite{dumbo}

@@

Most Hadoop users write their MapReduce programs in Java, since this is the
native language of Hadoop. \cite{java}

@@

Python is a popular general-purpose high-level multi-paradigm---most
notably: object-oriented, imperative and functional---scripting language.
\cite{python}

Zohmg is written mostly in Python. Zohmg uses the following Python modules
internally; it depends on them to be available on the system.

@@

Python Paste is a framework for building WSGI\footnote{Web Server Gateway
Interface, a HTTP extension where a CGI-like environment is passed around.}
applications and middleware. \cite{definitive_guide_to_pylons}

@@

Middleware is software that acts as an intermediary. WSGI middleware
receives a WSGI request and then performs logic upon this request, before
passing the request to a WSGI application. \cite{paste}

@@

PyYAML is a complete YAML 1.1 compliant parser and emitter for Python.
\cite{pyyaml}

@@

Python Setuptools is a framework---built on top of the Python module
distutils---for basic package management of Python software. Amongst other
things it creates and installs Python Eggs\footnote{A Zip file which
bundles Python software files.}. \cite{setuptools}

@@

A simple, fast, complete and extensible JSON encoder and decoder for
Python. The Python module simplejson is implemented entirely in Python and
has no external dependencies. However a C extension which provides a
serious speed boost is included. \cite{simplejson}

@@

\subsection{Data sets}

The user's data sets are defined using YAML. The data set files reside in
the config directory.

@@

The Darling output format for HBase. Between Zohmg and Hadoop JSON is used
for serializing data.

@@

The data export server is uilt with Python Paste---a middleware
framework---the data server dispatches to the requested application
depending on the URI it was fed.

Middleware which dispatches to applications based on the URI. The URLParser
removes the first part of the URI and tries to find a matching application
to relay the request to. For instance, consider the URI
\texttt{/client/file.html}. The first part---\texttt{client}---will be
removed and the request \texttt{/file.html} will be sent to the application
named \texttt{client.py}.

Dispatches request to serve static files from the project's
\texttt{clients} directory. Custom clients---which interact with Zohmg---can
be built with HTML and JavaScript, see the examples.

Fetches data from HBase via its Thrift interface. The data is served to the
user as JSONP\footnote{JSON with padding: JSON extension which adds name of
callback function as an argument to the function itself.}.

@@

The data server have three methods of serving data: raw, transformed, and
static. The raw and transformed data are aggregates served from HBase,
transformed in the latter case, while the static method just serve static
files from the project's \textit{clients} directory. Aggregates served from
HBase are returned to the user as JSONP.

\subsection{Data}

Based on the query---which the user submits---the corresponding data is
served from HBase as JSONP. The query is a simple definition of wanted row
range, dimensions, and values for these dimensions. The query is submitted
to the data server via HTTP, an example query could look like (the URL is
wrapped to fit the page) \\

\texttt{http://localhost:8086/data/?t0=20090101 \\
\&t1=20090420\&unit=pageviews\&d0=country\&d0v=SE} \\

The above query would return the pageviews for SE between 1 January 2009
and 4 April 2009.

\subsection{Transformers}

Fetches data from HBase and transforms it with the requested transform
program from the project's \texttt{transformers} directory, output is
dumped as JSONP.

\subsection{Clients}

It is possible to build static web pages out of HTML and JavaScript, which
can interact with the data server. These static pages can for instance
present some sort of query building mechanism (for insntance with drop down
boxes) and then send this query to the data server and graph the resulting
data.

The data server serves static files from the project's \texttt{clients}
directory at the server URL appended with \texttt{client/filename}.

@@

Configure the dataset.yaml.

Write a custom mapper, which imports data from web logs.

Start data server and extract data from it.

Possibly write a transformer which performs some transformation on the
aggregated data.

Possibly write a client which will render a graph.

@@

\subsection{Multiple data sets}

@@

\subsection{Updating data set files}

Update tables on changes in configuration---data set files.

@@

Last.fm is looking at using Zohmg for multiple application areas, such as
radio stats and web page stats.
